---
title: "Linear Models: Homework 1"
author: "Juan Andres VANEGAS JADAN"
date: "2024-2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r include=FALSE}
library(dplyr)
library(skimr)
library(ggplot2)
```

# Answers to the questions

## Question 1a

Each ug increase of ADDF17 has the effect of increasing 1.0714mg/L on the mean of IgG1 levels.

## Question 1b

The 95% confidence interval lies between 0.44 and 1.7 mg/L, indicating that the effect is
likely posible, as 0 is not included. This interval suggest that 
for each ug of ADDF17, the mean of IgG1 would increase between between 0.44 and 1.7 mg/L
with 95% of probability. 
With a standard error of 0.31, the estimation of the effect of ADDF17
over IgG1 has a moderate variability.

## Question 1c

The effect of ADDF17 over IgG1 levels is statistically significant, with a p-value of 
0.001, This result suggests that ADDF17 concentration has a meaningful impact on IgG1 levels.

The p-value was calculated with a two-sided test $H1: \neq 0$, as there was 
no prior information on wheather ADDF17 would increase or decrease IgG1 levels.

## Question 1d

If we do not include the doses of 2ug, we obtain a lower value for the estimation,
equals to 0.8252. Furthermore with a p-value of 0.239, this estimation would not be significant.

The non significance can also be reiterated looking at the confidence interval. This would 
include values from -0.56 to 2.21, giving inconclusive information if the 
ADDF17 concentration would increase or decrease the IgG1 levels.

## Question 2

The simulations reveal that the 2ug doses samples play a key role for obtaining
significant estimations. When we include them, 71% of the simulations return
a significant estimations. In contrast, when we exclude them, only 20% of the 
simulations return significant estimations. One reason could be that the higher doses 
samples have a larger and more consistent effect.

\newpage

# Appendix with R code

## Question 1

```{r}
home_path <- Sys.getenv("HOME")
setwd(paste(home_path, "/academics/hasselt/linear-models", sep = ""))
load(file = "Data/mice.RData")
str(mice)

plot(mice$concentration, mice$IgG1)
```

Definition of the linear model.
```{r}
m <- lm(mice$IgG1 ~ mice$concentration)
summary(m)
```
Estimation and variance of the estimation give us:
```{r}
coef_est <- coef(summary(m))
beta1_hat <- coef_est["mice$concentration", "Estimate"]
sigma2_beta1 <- coef_est["mice$concentration", "Std. Error"]
print(paste("The parameter beta1 is estimated to: ", beta1_hat))
print(paste("and it has a std deviation of ", sigma2_beta1))
```
Computation of the confidence interval give us:
```{r}
lowerci <- beta1_hat - qt(0.975, df = nrow(mice) - 2) * sigma2_beta1
upperci <- beta1_hat + qt(0.975, df = nrow(mice) - 2) * sigma2_beta1
print(paste("The confidence intervale is beetween ", lowerci, " and ", upperci))
```
Computation of the p-value at two sides give us:
```{r}
t_observed <- beta1_hat / sigma2_beta1
p.value.twosided <- 2 * (1 - pt(abs(t_observed), df = nrow(mice) - 2))
print(paste("The p-value of the parameter estimation is of ", p.value.twosided))
```
If we evaluate the same data but excluding the dose of 2ug
```{r}
mice_filtered <- mice %>% filter(concentration != 2)
m_filtered <- lm(IgG1 ~ concentration, data = mice_filtered)
summary(m_filtered)
confint(m_filtered)
```


## Question 2

```{r}
set.seed(2678)

p1 <- p2 <- vector(length = 1000)
for (i in 1:1000) {
  conc2 <- mice$concentration
  mice2 <- data.frame(
    concentration = mice$concentration,
    IgG1 = 3 + 0.83 * conc2 + rnorm(62, sd = 0.9)
  )
  conc1 <- mice$concentration[mice$concentration < 2]
  mice1 <- data.frame(
    concentration = conc1,
    IgG1 = 3 + 0.83 * conc1 + rnorm(60, sd = 0.9)
  )

  m2 <- lm(IgG1 ~ concentration, data = mice2)
  p2[i] <- summary(m2)$coef[2, 4]

  m1 <- lm(IgG1 ~ concentration, data = mice1)
  p1[i] <- summary(m1)$coef[2, 4]
}

mean(p2 < 0.05)
mean(p1 < 0.05)
```