# Exercise Blood Pressure

In a small dose-finding study of a blood pressure reducing drug, 40 high blood 
pressure patients (systolic blood pressure at least 150 mmHg) were randomised 
over 4 concentrations of the active compound (arginine) in the drug: 0, 2, 5 
and 10 mg per day. The outcome is the systolic blood pressure reduction after 
2 months, measured in mmHg. 
## Libraries
```{r}
library(ggplot2)
```

## Reading data
```{r}
load("Data/BloodPressure.RData")
head(BloodPressure)
```
## Plot
```{r}
ggplot(BloodPressure, aes(x = dose, y = bp.reduction)) +
  geom_point(color = "blue") +
  xlab("Drug dose") +
  ylab("Blood Pressure Reduction")
```


```{r}
m.bloodpressure <- lm(BloodPressure$bp.reduction ~ BloodPressure$dose)
```

```{r}
ggplot(
  BloodPressure,
  aes(x = dose, y = bp.reduction)
) +
  geom_point(color = "blue") +
  xlab("dose (mg / day") +
  ylab("systolic blood pressure reduction (mmHg)") +
  theme(
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 15),
  ) +
  geom_abline(
    intercept = m.bloodpressure$coefficients[1],
    slope = m.bloodpressure$coefficients[2],
  )
```

