# Simulation Study
Repeat the previous simulation study, but now with other values of the regressor 
and other numbers of replicates for each regressor value. In particular, consider 
the setting with only two different values of the regressor ( $x = 165$ and
$x = 185$) and at each of these values, consider 25 replicates. This makes a 
total of $n = 50$ observations, just like in the simulation study.

Check whether the LSEs of $\beta_0$ and $\beta_1$ are unbiased and compute the 
empirical variance of the estimators from the simulation study. How do these 
variances compare to the variances from the previous simulation study? Can you 
give an explanation?

```{r}
x <- c(165, 185)
x <- rep(x, 25)
N <- 1000

betaHat50 <- data.frame(beta0Hat = NA, beta1Hat = NA)

for (i in 1:N) {
  y <- 90 + 0.5 * x + rnorm(length(x), sd = 5)
  m <- lm(y ~ x)
  betaHat50[i, ] <- coef(m)
}
print(paste("Variance by simultaion: "))
print(var(betaHat50))
X <- matrix(nrow = 50, ncol = 2)
X[, 1] <- 1
X[, 2] <- x
Var50 <- solve(t(X) %*% X) * 25
print("Variance by formula:")
print(Var50)
```
```{r}
hist(betaHat50$beta1Hat, main = "n=50", xlab = "beta1-hat")
abline(v = 0.5, col = 2)
```