# 2.3.2 BLUE

Repeat this simulation study with 10000 Monte Carlo simulation runs, but now 
with another estimator of $\beta$. With our conventional notation, define the 
following esimator,
$$
\tilde{\beta} = ({X}^t{X}+d{I}_2)^{-1}{X}^t{Y},
$$

```{r}
x <- c(165, 170, 175, 180, 185) # five father's heights
X <- cbind(1, x) # design matrix
betaHatLSE <- data.frame(beta0Hat = NA, beta1Hat = NA)
betaHatNew <- data.frame(beta0Hat = NA, beta1Hat = NA)
MSELSE <- matrix(nrow = N, ncol = 2)
MSENew <- matrix(nrow = N, ncol = 2)
for (i in 1:N) {
  y <- 90 + 0.5 * x + rnorm(5, sd = 5) # random sample of 5 outcomes
  m <- lm(y ~ x)
  betaHatLSE[i, ] <- coef(m)
  MSELSE[i, ] <- (coef(m) - c(90, 0.5))^2
  betaTilde <- solve(t(X) %*% X + 0.005 * diag(2)) %*% t(X) %*% y
  betaHatNew[i, ] <- betaTilde
  MSENew[i, ] <- (betaTilde - c(90, 0.5))^2
}

print(colMeans(betaHatLSE))
print(colMeans(betaHatNew))
```

## Note about solve:
When used to solve a system of linear equations of the form (Ax = b), solve 
finds the vector (x) that satisfies the equation.
```{r}
A <- matrix(c(2, 1, 1, 3), nrow = 2)
b <- c(1, 2)
x <- solve(A, b)
print(x) 
`````
Here solve function finds the inverse of a matrix. When used with a 
single argument, solve computes the inverse of the matrix. This is equivalent 
to solving the equation (AX = I) for (X), where (I) is the identity matrix. 
```{r}
A <- matrix(c(2, 1, 1, 3), nrow = 2)
A_inv <- solve(A)
print(A_inv)
```