# Muscle mass
Scientists suspect that the muscle mass of people starts declining from a 
certain age onwards.  To verify this research question, a nutritionist 
randomly sampled 59 women, aged between 41 and 78. For these women; also the 
muscle mass was measured (we actually only have a proxy based on 
bioelectrical impedance measurements).

Perform a regression analysis and formulate an answer to this research 
question (including parameter estimates, confidence interval and hypothesis 
test). You may use the next chunk of R code for reading the 
data.

```{r}
muscles <- read.csv("Data/Muscles.txt", sep = " ")
names(muscles) <- c("muscle.mass", "age")
skim(muscles)
```

```{r}
m <- lm(muscles$muscle.mass ~ muscles$age)
m_summary <- summary(m)
beta1.hat <- coef(m_summary)[, "Estimate"]["muscles$age"]
sigmab1sq.hat <- coef(m_summary)[, "Std. Error"]["muscles$age"]
t.observed <- beta1.hat / sigmab1sq.hat
p.value.twosided <- 2 * (1 - pt(abs(t.observed), nrow(muscles) - 2))
ci <- c(
  beta1.hat - qt(0.975, nrow(muscles) - 2) * sigmab1sq.hat,
  beta1.hat + qt(0.975, nrow(muscles) - 2) * sigmab1sq.hat
)
```