
# Muscle mass
Scientists suspect that the muscle mass of people starts declining from a 
certain age onwards.  To verify this research question, a nutritionist 
randomly sampled 59 women, aged between 41 and 78. For these women; also the 
muscle mass was measured (we actually only have a proxy based on 
bioelectrical impedance measurements).

Perform a regression analysis and formulate an answer to this research 
question (including parameter estimates, confidence interval and hypothesis 
test). You may use the next chunk of R code for reading the 
data.

```{r}
muscles <- read.csv("Data/Muscles.txt", sep = " ")
names(muscles) <- c("muscle.mass", "age")
skim(muscles)
```
## Model
```{r}
m <- lm(muscle.mass ~ age, data = muscles)
```

## Linearity of the regression model
```{r}
e <- m$residuals
x.all <- unique(muscles$age)
ave.e <- c()
for (x in x.all) {
  ave.e <- c(ave.e, mean(e[muscles$age == x]))
}
par(mfrow=c(1,1))
plot(muscles$age, e,
  cex.lab = 1.5, cex.axis = 1.5,
  xlab = "Age (years)", ylab = "residuals"
)
points(x.all, ave.e, col = 2, pch = 8)
abline(h = 0, lty = 2)
```

## Normality of the error terms
```{r}
par(mfrow=c(1,2))
qqnorm(m$resid,cex.lab=1.5,cex.axis=1.5,
       xlab="expected quantiles",ylab="residuals",main="")
qqline(m$resid)

boxplot(m$residuals, ylab="residuals")
```

## Homoskedasticity
```{r}
par(mfrow=c(1,2))
e<-m$residuals
plot(muscles$age,e^2,cex.lab=1.5,cex.axis=1.5,
     xlab="Age (years)",
     ylab="Squared residuals")
abline(h=sum(m$residuals^2)/(173-2),lty=2)

e<-m$residuals
plot(muscles$age,abs(e),cex.lab=1.5,cex.axis=1.5,
     xlab="Age (years)"
     ,ylab="absolute value of the residuals")
abline(h=sum(abs(m$residuals))/(173-2),lty=2)
```

