## Model Assesment
## Libraries
```{r}
library(dplyr)
library(skimr)
library(ggplot2)
```
## Load data
```{r}
home_path <- Sys.getenv("HOME")
setwd(paste(home_path, "/academics/hasselt/linear-models", sep = ""))
Galton <- read.csv("./Data/galton.tab", sep = "\t")
```
## Filter
```{r}
Galton.sons <- Galton %>%
  filter(gender == "M") %>%
  group_by(family) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
    father.cm = father * 2.54,
    son.cm = height * 2.54
  ) %>%
  select(father.cm, son.cm)
```

## Linearity of the regression model
```{r}
m <- lm(son.cm ~ father.cm, data = Galton.sons)
e <- m$residuals
x.all <- unique(Galton.sons$father.cm)
ave.e <- c()
for (x in x.all) {
  ave.e <- c(ave.e, mean(e[Galton.sons$father.cm == x]))
}
plot(Galton.sons$father.cm, e,
  cex.lab = 1.5, cex.axis = 1.5,
  xlab = "Father's height (cm)", ylab = "residuals"
)
points(x.all, ave.e, col = 2, pch = 8)
abline(h = 0, lty = 2)
```
- doesn't show any systemathic pattern

## Normality of the error term
```{r}
qqnorm(m$resid,
  cex.lab = 1.5, cex.axis = 1.5,
  xlab = "expected quantiles", ylab = "residuals", main = ""
)
qqline(m$resid)
```
Figure shows the normal QQ-plot of the residuals of the Galton example. Most 
of the points in the QQ-plot are close to the straight line with no 
systematic pattern, and with only a few larger deviations in the right hand 
tail of the distribution. However, the number of outliers (2 or 3) is very 
small as compared to the sample size (173). The plot does also not reveal 
systematic deviations from the straight line.

## Homoskedasticity

- Plots $e^2$
```{r}
par(mfrow = c(1, 2))
heights <- unique(Galton.sons$father.cm)
var.y <- c()
for (x in heights) {
  # var.y<-c(var.y,var(Galton.sons$son.cm[Galton.sons$father.cm==x]))
  var.y <- c(var.y, mean(m$residuals[Galton.sons$father.cm == x]^2))
}
plot(heights, var.y,
  cex.lab = 1.5, cex.axis = 1.5,
  xlab = "Father's height (cm)",
  ylab = "sample variance", col = 2, pch = 8
)
abline(h = sum(m$residuals^2) / 173, lty = 2)
# note that the reference line is at MSE (n-2)/n

plot(heights, sqrt(var.y),
  cex.lab = 1.5, cex.axis = 1.5,
  xlab = "Father's height (cm)",
  ylab = "sample standard deviation",
  col = 2, pch = 8
)
abline(h = sqrt(sum(m$residuals^2) / 173), lty = 2)
```
- Plots $|e|$
```{r}
par(mfrow = c(1, 2))
e <- m$residuals
plot(Galton.sons$father.cm, e^2,
  cex.lab = 1.5, cex.axis = 1.5,
  xlab = "Father's height (cm)",
  ylab = "Squared residuals"
)
abline(h = sum(m$residuals^2) / (173 - 2), lty = 2)

e <- m$residuals
plot(Galton.sons$father.cm, abs(e),
  cex.lab = 1.5, cex.axis = 1.5,
  xlab = "Father's height (cm)",
  ylab = "absolute value of the residuals"
)
abline(h = sum(abs(m$residuals)) / 173, lty = 2)
```